WSClean: an implementation of a fast, generic wide-field imager for radio astronomy

### 2014-03-31 articles/2014/Article-WSClean/wsclean.tex
Comments by Bryan G

\begin{abstract}
Widefield interferometric imaging of radio data with existing imager implementations is computationally expensive, especially for the large data volumes created by modern non-coplanar many-element arrays. We present a new widefield interferometric imager that uses the $w$-stacking algorithm and can make use of the $w$-snapshot algorithm. The performance dependencies of \textsc{casa}'s $w$-projection and our new imager are analysed and analytical functions are derived that describe the required computing cost for both imagers. On data from the Murchison Widefield Array, we find our new method to be an order of magnitude faster than $w$-projection, as well as being capable of full-sky imaging at full resolution and with correct polarization correction.  We predict the computing costs for several other arrays and estimate that our imager is a factor of 2--12 faster, depending on the array configuration. We estimate the computing cost for imaging the low-frequency Square-Kilometre Array observations to be 60 PetaFLOPS with current techniques. We find that combining $w$-stacking with the $w$-snapshot algorithm does not significantly improve computing requirements over pure $w$-stacking. The source code of our new imager is publicly released.
\end{abstract}

@@

\section{Introduction}
Visibility data from non-coplanar interferometric radio telescopes which observe large fractions of the sky at once can not be accurately imaged with a two-dimensional fast Fourier transform (FFT). Instead, the imaging algorithm needs to account for the ``$w$-term'' during inversion, which is the term that describes the deviation of the array from a perfect plane \citep{perley-noncoplanar-arrays}. The image degradation effects of the $w$-term are amplified for telescopes with wide fields of view (FOV), making this a significant issue for low-frequency telescopes that by nature are wide-field instruments.

@@

An interferometer samples the complex visibility function
\begin{align}\notag
V(u,v,w) = & \iint \frac{A(l,m) I(l,m)}{\sqrt{1-l^2-m^2}} \times \\ \label{eq:visibility-function}
& e^{-2\pi i \left(ul + vm + w(\sqrt{1-l^2-m^2}-1)\right)} dl dm,
\end{align}
where $u,v,w$ is a baseline coordinate in the coordinate system of the array, $A$ is the primary-beam function, $I$ is the sky function and $l,m$ are cosine sky coordinates. We will use $I'(l,m)$ to denote the sky function before primary-beam correction, $I'(l,m)=A(l,m)I(l,m)$. We will not discuss calibration, but assume $V$ has been calibrated before imaging. In the case of a polarized measurement, the symbols become $2\times 2$ matrices and beam correction is more complicated, but without loss of generality we will ignore polarization and treat inversion as a scalar problem. Imaging consists of inverting Eq.~\eqref{eq:visibility-function}, i.e., to find $I'$ from $V$.

@@

In practice, the final scaling will be different from $\left(w_{\max} - w_{\min}\right)/\sqrt{1-l^2-m^2}$ suggested by Eq.~\eqref{eq:wstacking}, because the individual $w$-layers will not be completely filled with samples. Therefore, the number of added samples and their weight have to be taken into account. Therefore, each pixel is divided by the weighted number of samples. Additionally, it might be required to divide out the effect of a possible convolution kernel, the primary beam and correct for other direction-dependent effects. In XX, YY, LL or RR polarizations, a correlated baseline is the complex conjugate of the reversed baseline, and the relation $V(u,v,w)=\overline{V(-u,-v,-w)}$ holds. The right-hand side of Eq.\eqref{eq:wstacking} with only positive $w$-value samples then becomes the complex conjugate of the one with only negative $w$-value samples. In this case we can therefore calculate the image for $w<0$ from the image with $w>0$. That allows us to set $w_{\min}$ to the minimum absolute $w$-value, which requires half the number of layers. In any case, the input to the two-dimensional inverse FFT is not generally an Hermitian-symmetric function, and hence the inverse FFT is always performed as a complex-to-complex transform.

@@

\subsection{Discretization of $w$} \label{sec:gridding-w}
\begin{figure}
\begin{center}
\includegraphics[width=8cm]{img/aliasing-example-annotated}
\caption{Aliasing artefacts caused by insufficient $w$-layers in a simulated field. \textsc{wsclean} was set to use 12 $w$-layers. The centre of the image is at 10\degree zenith angle, which would normally require $\sim195$ $w$-layers. Sources are 1~Jy (red circles), ghost sources are approximately 0.2~Jy. Each source produces two ghost sources, but because they re-appear after a major cleaning cycle, they are eventually cleaned and produce more ghost sources. }
\label{fig:aliasing-example}
\end{center}
\end{figure}
While the discretization of $u$ and $v$ is similar to conventional imaging, the discretization of $w$ defines the number of $w$-layers that need to be processed. For this, one can use a rule similar to \eqref{eq:when-2d-is-valid}, and make sure that the phase difference for two subsequent discretized $w$-values, $w_A$ and $w_B$,  is less than one radian. This results in the constraint
\begin{equation} \label{eq:minimum-w-distance}
\left|\left(w_A - w_B\right) 2\pi (\sqrt{1-l^2-m^2}-1)\right| \ll 1.
\end{equation}
This suggests that a uniform discretization in $w$ is optimal. This is in contrast to \citet{wprojection-cornwell} where $\sqrt{w}$ tabulation is suggested. From Eq.~\eqref{eq:minimum-w-distance}, the required number of layers can be derived and is given by
\begin{equation} \label{eq:nwlayers-bound}
 N_\textrm{wlay} \gg 2\pi \left(w_{\max} - w_{\min}\right) \max_{l,m} \left(1 - \sqrt{1-l^2-m^2}\right).
\end{equation}
Actual values for the right-hand side can be very different, depending on the observation. The value of $w_{\max} - w_{\min}$ is influenced by the coplanarity of the array, the zenith angle (ZA) and the wavelength, while the value of the $\max_{l,m}$ term is influenced by the angular size of the image. For the MWA, a typical value of $w_{\max} - w_{\min}$ is $\sim 10$ at zenith, and reaches $\sim 400$ at a ZA of $30^{\circ}$. For a typical full-field-of-view image of a MWA observation of $3072\times 3072$ pixels of $0.75'$ size, the $\max_{l,m}$ term is $0.68$. This implies that tens of $w$-layers are required at zenith and hundreds at lower elevations. The number of $w$-layers has a large effect on the performance of the $w$-stacking algorithm, and will be discussed further in the next sections.

@@

\section{The \textsc{wsclean} imager implementation} \label{sec:implementation}
With the purpose of testing new algorithms for the imaging of MWA data, we have written a new imager around the $w$-stacking algorithm. The imager is not specialized for the MWA, and has been successfully used for imaging VLA and GMRT data.

The new imager, called ``\textsc{wsclean}'', is written in the C++ language. It reads visibilities from \textsc{casa} measurement sets and writes output images to Flexible Image Transport System (FITS) files. Several steps are multi-threaded using the threading module of the C++11 standard library. These are: reading and writing; gridding from and to different $w$-layers; performing the FFTs; and H\"ogbom Clean iterations (peak-finding and image subtraction, \citealt{hogbom-clean}). For the latter, intrinsics are used as well. Because of these optimizations, minor Cleaning iterations with an image size of $3072\times3072$ are performed at a rate of hundreds per second. This is fast enough to make the Clark Clean optimization \citep{clark-clean}, which consists of considering only a subset of pixels with a reduced point-spread function (PSF), less relevant. For very large images this optimization might still be useful, but we have not implemented it in \textsc{wsclean}. Because the PSF varies for imaging with non-zero $w$-values, subtracting a constant PSF in image space leads to inaccuracies. After a number of minor iterations, it is therefore beneficial to invert the model back to the visibilities via prediction, and subtract the model directly from the visibilities \citep{wprojection-cornwell}. This is similar to the Cotton-Schwab cleaning method \citep{cotton-schwab-clean}. \textsc{wsclean} allows cleaning individual polarizations, or can jointly deconvolve the polarizations. In the latter case, peaks are searched in $pp^2+2pq \overline{(pq)} + qq^2$ space, where $p$ and $q$ are the two polarizations and $\overline{pq}=qp$ is the complex conjugate of $pq$. After a peak has been found, the PSF is subtracted from the individual polarizations with different factors.

@@

\begin{figure*}
\begin{center}
\includegraphics[height=22.5cm]{img/fullsky-together}
\caption{A beam-corrected MWA image of a 112~s zenith observation at 180~MHz that covers almost the full sky. The lower image shows a zoom-in on the southern side lobe. PKS J2358-6054 ($\sim100$ Jy) is visible in the centre of the southern side lobe and resolved. The noise level in the side lobe is about 200 mJy. PKS J2358-6054 has been cleaned, but some artefacts remain because no direction-dependent calibration has been performed.}
\label{fig:full-sky-example}
\end{center}
\end{figure*}

@@

\subsection{$w$-snapshots} \label{sec:snapshot-imaging}
\begin{figure*}
\begin{center}
\includegraphics[width=8cm]{img/vela-normal-projection}
\includegraphics[width=8cm]{img/vela-zenith-projection}
\caption{13-min MWA observation of supernova remnants Vela and Puppis A. Left: normal projection centred on Puppis A, right: phase-rotated to zenith to reduce $w$-values, recentred on Puppis A during imaging. Both Stokes-I images have been made with \textsc{wsclean} using the Cotton-Schwab clean algorithm. No beam correction was applied. Imaging computing cost was 60 min for the normal projected image and 41 min for the recentred image.}
\label{fig:vela-projection-example}
\end{center}
\end{figure*}
As discussed, all-sky snapshot imaging with zenith as phase centre is quite efficient. However, if one is not interested in imaging the whole sky, and the direction of interest is far from zenith, the computational overhead of making all-sky images is undesirable.

Instead of imaging the whole sky, the image can be recentred from $(l,m)$ to $(\hat{l},\hat{m})=(l+\Delta l,m+\Delta m)$ by performing the substitution $(l,m)\leftarrow(\hat{l},\hat{m})$ in Eq.~\eqref{eq:wstacking}:
\begin{align}\notag
\frac{I'(\hat{l}, \hat{m})\left(w_{\max} - w_{\min}\right)}{\sqrt{1-\hat{l}^2-\hat{m}^2}} = \int\limits_{w_{\min}}^{w_{\max}} e^{2\pi i w(\sqrt{1-\hat{l}^2-\hat{m}^2}-1)} \times \\
\iint e^{2 \pi i \left(u\Delta l + v\Delta m\right)} V(u,v,w) e^{2\pi i \left(ul + vm\right)} du dv dw.
\end{align}
In words, recentering an image involves accounting for the position shift during the $w$-correction and final scaling, and shifting the visibilities in phase by multiplication with $e^{2 \pi i \left(u\Delta l + v\Delta m\right)}$ prior to the imaging. This essentially is a method of implementing the $w$-snapshots algorithm as described by \citet{widefield-imaging-ska-cornwell}. However, instead of creating warped images by performing the FFT over a tilted plane in $uvw$-space, our method transforms planes with constant $w$-values. The benefit of our method is that the resulting image has a circular synthesized beam, i.e., the resolution in $l$ and $m$ directions matches the intrinsic resolution of the instrument, which is desirable for cleaning. Regridding a recentred image is also more straightforward compared to regridding a warped snapshot, because in the latter case there is no analytic solution to the coordinate conversion \citep{perley-noncoplanar-arrays}. A benefit of warped snapshots is that the $w$-term errors are zero at image centre and get worse with the distance from image centre. With our approach, the effect gets worse with the distance from zenith. Our technique has a similar computational cost compared to the $w$-snapshot algorithm, although the required number of $w$-layers has a different dependency on ZA, FOV and the non-coplanarity.

By doing the inverse corrections during prediction, a recentred visibility set can be cleaned with Cotton-Schwab iterations similar to a non-recentred set. The PSF used during minor cleaning iterations is created by multiplying the weights with the recentering corrections, such that the PSF represents a source in the middle of the image.

A recentred image is in the projection of the zenith tangent plane and, like warped snapshots, will need an additional regridding step before multiple snapshots can be added together. A recentred image can be stored in a FITS file with the orthographic (SIN) projection normally used in interferometric imaging \citep{fits-coordinates-2002}, by setting the centre of tangent projection with the \texttt{CRPIXi} keyword \citep{wcs-in-fits}. Common viewers such as \textsc{kvis} \citep{karma-1996} and \textsc{ds9} support such FITS files and display their coordinates correctly. Unlike warped snapshots, recentred images do not need the generalized-SIN-projection keywords \texttt{PV2\_1} and \texttt{PV2\_2}.

An example of the difference in projection between non-recentred and recentred imaging is given in Fig.~\ref{fig:vela-projection-example}, which displays a 13-min MWA observation imaged with \textsc{wsclean}. The processing steps performed for this image were: i) preprocessing of the observation with the Cotter preprocessing pipeline, which includes time averaging and RFI flagging with the \textsc{aoflagger} \citep{post-correlation-rfi-classification,scale-invariant-rank-operator}; ii) calibration using Hydra~A without direction dependence using a custom implementation of the RTS full-polarization calibration algorithm \citep{rts-mwa}; and iii) imaging of the seven 112 s intervals separately. Cleaning an image that is in a different projection yields slightly different results, but qualitatively the two images are clearly of equal accuracy. The wall-clock time for imaging is 60 min and 41 min for normal projection and the recentred image, respectively. Of the 41 min, 3 min is spent on phase rotating the visibilities.

Another method to shift an image is by using the periodicity of the FFT function. Since sources outside the FOV will be aliased back into the field, the image can be transformed to have the alias in the centre. This complicates gridding during imaging, because the anti-aliasing kernel needs to have a pass-band shape instead of a low-pass shape. Therefore, we chose to implement the former method.

\subsection{Beam correction}
\begin{figure*}
\begin{center}
\includegraphics[width=6cm]{img/pulsar-stokesi}\hspace{1cm}\includegraphics[width=6cm]{img/pulsar-stokesv}
\caption{Stokes I (left panel, total power) and Stokes V (right panel, circular polarization) images of PSR J0437-4715, demonstrating the full-polarization beam correction capability for MWA observations with \textsc{wsclean}. This observation was performed in drift-scan mode. The pulsar (centre of image) displays 17\% circular polarization. Due to inaccuracies in the current beam model, other (unpolarized) sources show $\sim$1\% leakage into Stokes~V.}
\label{fig:pulsar-stokes-iv}
\end{center}
\end{figure*}
Because the MWA consists of fixed, beam-formed dipole antennas, the voltage beam of a MWA tile is described by a complex, non-diagonal Jones matrix. Alternatively, Mueller matrices can be used as well, but in general these result in more computations \citep{revisiting-me-i}. When assuming all individual antennas of the tiles are working properly, all tiles have the same beam. The beam can then be corrected in image space for all tiles at once with
\begin{equation}\label{eq:beam-correction}
 I(l, m) = J(l,m)^{-1} I'(l, m) \left(J(l,m)^*\right)^{-1},
\end{equation}
where each term is a $2\times2$ complex matrix, with $J$ the voltage beam and $^*$ denoting the conjugate transpose. Because $J$ is complex and non-diagonal, all four combinations of the cross-correlated polarizations $p$ and $q$\footnote{The instrumental polarizations of the MWA are informally often referred to as $X$ and $Y$, mainly because many software treats the two polarizations as such. However, this is somewhat confusing because they are not necessarily orthogonal.} including the imaginary part of the $pq$ and $qp$ are required to calculate any of the Stokes parameters. The $pp$ and $qq$ have no imaginary part due to the $uv$-symmetry. To make proper MWA beam correction possible, \textsc{wsclean} can output both real $pp$, $pq$, $qp$, $qq$ images and imaginary $pq$ and $qp$ images. This requires four runs of the algorithm. Once these images are created, a separate program is used to perform the correction of Eq.~\eqref{eq:beam-correction}. The method is demonstrated in Fig.~\ref{fig:pulsar-stokes-iv}, which displays a detection of pulsar J0437-4715 in Stokes V. The image was made with an initial pipeline for the MWA radio-sky monitor project (PI: Bell) that uses the MWA to search for transient and variable sources. The pipeline includes \textsc{wsclean} to make the Stokes-parameter images.

@@

\subsection{Decreasing inversion resolution} \label{sec:decreasing-inversion-resolution}
Typically, for cleaning it is desired that images have a pixel size (side length) at least five times smaller than the synthesized-beam width. This improves the accuracy of the clean algorithm, because the positions of image maxima will be closer to the actual source positions. This factor of five is normally taken into account in the overall imaging resolution, i.e., the image size is increased during inversion. For the $w$-projection method, the cost of inversion with an increased resolution is small, because it does not increase the size of the $w$-kernels. It will affect the inverse FFT, but the relative cost of this step is negligible in the total cost. The $w$-stacking method is however significantly affected by the image resolution, because it performs many inverse FFTs.

The spatial frequencies in the output image are band-limited by the synthesized beam. Therefore, it is not required to perform the inversion at higher resolution. As long as the $uv$-plane is sampled with at least the Nyquist frequency, a high resolution image can be perfectly reconstructed from an inversion at lower resolution. After an enlarged image has been cleaned, a model is created at the same (high) resolution. Because the model consists of delta functions, the spatial frequencies in the model are not band-limited. Consequently, decimating the model image will remove information from the model. However, only the low spatial frequencies of this image will be used in the prediction. Therefore, inversion at higher resolution is not required: as long as the decimation does not modify the low-frequency components, the output of the prediction step will not change. This requires low-pass filtering of the model before decimation.

>>>>

\begin{abstract}
Astronomical widefield imaging of interferometric radio data is computationally expensive, especially for the large data volumes created by modern non-coplanar many-element arrays. We present a new widefield interferometric imager that uses the $w$-stacking algorithm and can make use of the $w$-snapshot algorithm. The performance dependencies of \textsc{casa}'s $w$-projection and our new imager are analysed and analytical functions are derived that describe the required computing cost for both imagers. On data from the Murchison Widefield Array, we find our new method to be an order of magnitude faster than $w$-projection, as well as being capable of full-sky imaging at full resolution and with correct polarisation correction.  We predict the computing costs for several other arrays and estimate that our imager is a factor of 2--12 faster, depending on the array configuration. We estimate the computing cost for imaging the low-frequency Square-Kilometre Array observations to be 60 PetaFLOPS with current techniques. We find that combining $w$-stacking with the $w$-snapshot algorithm does not significantly improve computing requirements over pure $w$-stacking. The source code of our new imager is publicly released.
\end{abstract}

@@

\section{Introduction}
Visibility data from non-coplanar interferometric radio telescopes that observe large fractions of the sky at once can not be accurately imaged with a two-dimensional fast Fourier transform (FFT). Instead, the imaging algorithm needs to account for the ``$w$-term'' during inversion, which is the term that describes the deviation of the array from a perfect plane \citep{perley-noncoplanar-arrays}. The image degradation effects of the $w$-term are amplified for telescopes with wide fields of view (FOV), making this a significant issue for low-frequency telescopes that by nature are wide-field instruments.

@@

An interferometer samples the complex visibility function
\begin{align}\notag
V(u,v,w) = & \iint \frac{A(l,m) I(l,m)}{\sqrt{1-l^2-m^2}} \times \\ \label{eq:visibility-function}
& e^{-2\pi i \left(ul + vm + w(\sqrt{1-l^2-m^2}-1)\right)} dl dm,
\end{align}
where $u,v,w$ is a baseline coordinate in the coordinate system of the array, $A$ is the primary-beam function, $I$ is the sky function and $l,m$ are cosine sky coordinates. We will use $I'(l,m)$ to denote the sky function before primary-beam correction, $I'(l,m)=A(l,m)I(l,m)$. We will not discuss calibration, but assume $V$ has been calibrated before imaging. In the case of a polarised measurement, the symbols become $2\times 2$ matrices and beam correction is more complicated, but without loss of generality we will ignore polarisation and treat inversion as a scalar problem. Imaging consists of inverting Eq.~\eqref{eq:visibility-function}, i.e., to find $I'$ from $V$.

@@

In practice, the final scaling will be different from $\left(w_{\max} - w_{\min}\right)/\sqrt{1-l^2-m^2}$ suggested by Eq.~\eqref{eq:wstacking}, because the individual $w$-layers will not be completely filled with samples. Therefore, the number of added samples and their weights have to be taken into account. Therefore, each pixel is divided by the weighted number of samples. Additionally, it might be required to divide out the effect of a possible convolution kernel, the primary beam and correct for other direction-dependent effects. In XX, YY, LL or RR polarisations, a correlated baseline is the complex conjugate of the reversed baseline, and the relation $V(u,v,w)=\overline{V(-u,-v,-w)}$ holds. The right-hand side of Eq.\eqref{eq:wstacking} with only positive $w$-value samples then becomes the complex conjugate of the one with only negative $w$-value samples. In this case we can therefore calculate the image for $w<0$ from the image with $w>0$. That allows us to set $w_{\min}$ to the minimum absolute $w$-value, which requires half the number of layers. In any case, the input to the two-dimensional inverse FFT is not generally a Hermitian-symmetric function, and hence the inverse FFT is always performed as a complex-to-complex transform.

@@

\subsection{Discretisation of $w$} \label{sec:gridding-w}
\begin{figure}
\begin{center}
\includegraphics[width=8cm]{img/aliasing-example-annotated}
\caption{Aliasing artefacts caused by insufficient $w$-layers in a simulated field. \textsc{wsclean} was set to use 12 $w$-layers. The centre of the image is at 10\degree zenith angle, which would normally require $\sim195$ $w$-layers. Sources are 1~Jy (red circles), ghost sources are approximately 0.2~Jy. Each source produces two ghost sources, but because they re-appear after a major cleaning cycle, they are eventually cleaned and produce more ghost sources. }
\label{fig:aliasing-example}
\end{center}
\end{figure}
While the discretisation of $u$ and $v$ is similar to conventional imaging, the discretisation of $w$ defines the number of $w$-layers that need to be processed. For this, one can use a rule similar to \eqref{eq:when-2d-is-valid}, and make sure that the phase difference for two subsequent discretised $w$-values, $w_A$ and $w_B$,  is less than one radian. This results in the constraint
\begin{equation} \label{eq:minimum-w-distance}
\left|\left(w_A - w_B\right) 2\pi (\sqrt{1-l^2-m^2}-1)\right| \ll 1.
\end{equation}
This suggests that a uniform discretisation in $w$ is optimal. This is in contrast to \citet{wprojection-cornwell} where $\sqrt{w}$ tabulation is suggested. From Eq.~\eqref{eq:minimum-w-distance}, the required number of layers can be derived and is given by
\begin{equation} \label{eq:nwlayers-bound}
 N_\textrm{wlay} \gg 2\pi \left(w_{\max} - w_{\min}\right) \max_{l,m} \left(1 - \sqrt{1-l^2-m^2}\right).
\end{equation}
Actual values for the right-hand side can be very different, depending on the observation. The value of $w_{\max} - w_{\min}$ is influenced by the coplanarity of the array, the zenith angle (ZA) and the wavelength, while the value of the $\max_{l,m}$ term is influenced by the angular size of the image. For the MWA, a typical value of $w_{\max} - w_{\min}$ is $\sim 10$ at zenith, and reaches $\sim 400$ at a ZA of $30^{\circ}$. For a typical full-field-of-view image of a MWA observation of $3072\times 3072$ pixels of $0.75'$ size, the $\max_{l,m}$ term is $0.68$. This implies that tens of $w$-layers are required at zenith and hundreds at lower elevations. The number of $w$-layers has a large effect on the performance of the $w$-stacking algorithm, and will be discussed further in the next sections.

@@

\section{The \textsc{wsclean} imager implementation} \label{sec:implementation}
With the purpose of testing new algorithms for the imaging of MWA data, we have written a new imager around the $w$-stacking algorithm. The imager is not specialised for the MWA, and has been successfully used for imaging VLA and GMRT data.

The new imager, called ``\textsc{wsclean}'', is written in the C++ language. It reads visibilities from \textsc{casa} measurement sets and writes output images to Flexible Image Transport System (FITS) files. Several steps are multi-threaded using the threading module of the C++11 standard library. These are: reading and writing; gridding from and to different $w$-layers; performing the FFTs; and H\"ogbom Clean iterations (peak-finding and image subtraction, \citealt{hogbom-clean}). For the latter, intrinsics are used as well. Because of these optimisations, minor Cleaning iterations with an image size of $3072\times3072$ are performed at a rate of hundreds per second. This is fast enough to make the Clark Clean optimisation \citep{clark-clean}, which consists of considering only a subset of pixels with a reduced point-spread function (PSF), less relevant. For very large images this optimisation might still be useful, but we have not implemented it in \textsc{wsclean}. Because the PSF varies for imaging with non-zero $w$-values, subtracting a constant PSF in image space leads to inaccuracies. After a number of minor iterations, it is therefore beneficial to invert the model back to the visibilities via prediction, and subtract the model directly from the visibilities \citep{wprojection-cornwell}. This is similar to the Cotton-Schwab cleaning method \citep{cotton-schwab-clean}. \textsc{wsclean} allows cleaning individual polarisations, or can jointly deconvolve the polarisations. In the latter case, peaks are searched in $pp^2+2pq \overline{(pq)} + qq^2$ space, where $p$ and $q$ are the two polarisations and $\overline{pq}=qp$ is the complex conjugate of $pq$. After a peak has been found, the PSF is subtracted from the individual polarisations with different factors.

@@

\begin{figure*}
\begin{center}
\includegraphics[height=22.5cm]{img/fullsky-together}
\caption{A beam-corrected MWA image of a 112~s zenith observation at 180~MHz that covers almost the full sky. The lower image shows a zoom-in on the southern side lobe. PKS J2358-6054 ($\sim100$ Jy) is visible in the centre of the southern side lobe and resolved. The noise level in the side lobe is about 200 mJy/beam. PKS J2358-6054 has been cleaned, but some artefacts remain because no direction-dependent calibration has been performed.}
\label{fig:full-sky-example}
\end{center}
\end{figure*}

@@

\subsection{$w$-snapshots} \label{sec:snapshot-imaging}
\begin{figure*}
\begin{center}
\includegraphics[width=8cm]{img/vela-normal-projection}
\includegraphics[width=8cm]{img/vela-zenith-projection}
\caption{13-min MWA observation of supernova remnants Vela and Puppis A with a centre frequency of 149~MHz. Left: normal projection centred on Puppis~A, right: phase-rotated to zenith to reduce $w$-values, recentred on Puppis A during imaging. Both Stokes-I images have been made with \textsc{wsclean} using the Cotton-Schwab clean algorithm. No beam correction was applied. Imaging computing cost was 60 min for the normal projected image and 41 min for the recentred image.}
\label{fig:vela-projection-example}
\end{center}
\end{figure*}
As discussed, all-sky snapshot imaging with zenith as phase centre is quite efficient. However, if one is not interested in imaging the whole sky, and the direction of interest is far from zenith, the computational overhead of making all-sky images is undesirable.

Instead of imaging the whole sky, the image can be recentred from $(l,m)$ to $(\hat{l},\hat{m})=(l+\Delta l,m+\Delta m)$ by performing the substitution $(l,m)\rightarrow(\hat{l},\hat{m})$ in Eq.~\eqref{eq:wstacking}:
\begin{align}\notag
\frac{I'(\hat{l}, \hat{m})\left(w_{\max} - w_{\min}\right)}{\sqrt{1-\hat{l}^2-\hat{m}^2}} = \int\limits_{w_{\min}}^{w_{\max}} e^{2\pi i w(\sqrt{1-\hat{l}^2-\hat{m}^2}-1)} \times \\
\iint e^{2 \pi i \left(u\Delta l + v\Delta m\right)} V(u,v,w) e^{2\pi i \left(ul + vm\right)} du dv dw.
\end{align}
In words, recentring an image involves accounting for the position shift during the $w$-correction and final scaling, and shifting the visibilities in phase by multiplication with $e^{2 \pi i \left(u\Delta l + v\Delta m\right)}$ prior to the imaging. This essentially is a method of implementing the $w$-snapshots algorithm as described by \citet{widefield-imaging-ska-cornwell}. However, instead of creating warped images by performing the FFT over a tilted plane in $uvw$-space, our method transforms planes with constant $w$-values. The benefit of our method is that the resulting image has a circular synthesised beam, i.e., the resolution in $l$ and $m$ directions matches the intrinsic resolution of the instrument, which is desirable for cleaning. Regridding a recentred image is also more straightforward compared to regridding a warped snapshot, because in the latter case there is no analytic solution to the coordinate conversion \citep{perley-noncoplanar-arrays}. A benefit of warped snapshots is that the $w$-term errors are zero at image centre and get worse with the distance from image centre. With our approach, the effect gets worse with the distance from zenith. Our technique has a similar computational cost compared to the $w$-snapshot algorithm, although the required number of $w$-layers has a different dependency on ZA, FOV and the non-coplanarity.

By doing the inverse corrections during prediction, a recentred visibility set can be cleaned with Cotton-Schwab iterations similar to a non-recentred set. The PSF used during minor cleaning iterations is created by multiplying the weights with the recentring corrections, such that the PSF represents a source in the middle of the image.

A recentred image is in the projection of the zenith tangent plane and, like warped snapshots, will need an additional regridding step before multiple snapshots can be added together. A recentred image can be stored in a FITS file with the orthographic (SIN) projection normally used in interferometric imaging \citep{fits-coordinates-2002}, by setting the centre of tangent projection with the \texttt{CRPIXi} keyword \citep{wcs-in-fits}. Common viewers such as \textsc{kvis} \citep{karma-1996} and \textsc{ds9} support such FITS files and display their coordinates correctly. Unlike warped snapshots, recentred images do not need the generalised-SIN-projection keywords \texttt{PV2\_1} and \texttt{PV2\_2}.

An example of the difference in projection between non-recentred and recentred imaging is given in Fig.~\ref{fig:vela-projection-example}, which displays a 13-min MWA observation imaged with \textsc{wsclean}. The processing steps performed for this image were: i) preprocessing of the observation with the Cotter preprocessing pipeline, which includes time averaging and RFI flagging with the \textsc{aoflagger} \citep{post-correlation-rfi-classification,scale-invariant-rank-operator}; ii) calibration using Hydra~A without direction dependence using a custom implementation of the RTS full-polarisation calibration algorithm \citep{rts-mwa}; and iii) imaging of the seven 112 s intervals separately. Cleaning an image that is in a different projection yields slightly different results, but qualitatively the two images are clearly of equal accuracy. The wall-clock time for imaging is 60 min and 41 min for normal projection and the recentred image, respectively. Of the 41 min, 3 min is spent on phase rotating the visibilities.

Another method to shift an image is by using the periodicity of the FFT function. Since sources outside the FOV will be aliased back into the field, the image can be transformed to have the alias in the centre. This complicates gridding during imaging, because the anti-aliasing kernel needs to have a pass-band shape instead of a low-pass shape. Therefore, we chose to implement the former method of recentring the phase centre before imaging.

\subsection{Beam correction}
\begin{figure*}
\begin{center}
\includegraphics[width=6cm]{img/pulsar-stokesi}\hspace{1cm}\includegraphics[width=6cm]{img/pulsar-stokesv}
\caption{Stokes I (left panel, total power) and Stokes V (right panel, circular polarisation) images of PSR J0437-4715, demonstrating the full-polarisation beam correction capability for MWA observations with \textsc{wsclean}. This observation was performed in drift-scan mode with a centre frequency of 154~MHz. The pulsar (centre of image) displays 17\% circular polarisation. Due to inaccuracies in the current beam model, other (unpolarised) sources show $\sim$1\% leakage into Stokes~V.}
\label{fig:pulsar-stokes-iv}
\end{center}
\end{figure*}
Because the MWA consists of fixed, beam-formed dipole antennas, the voltage beam of a MWA tile is described by a complex, non-diagonal Jones matrix. Alternatively, Mueller matrices can be used as well, but in general these result in more computations \citep{revisiting-me-i}. When assuming all individual antennas of the tiles are working properly, all tiles have the same beam. The beam can then be corrected in image space for all tiles at once with
\begin{equation}\label{eq:beam-correction}
 I(l, m) = J(l,m)^{-1} I'(l, m) \left(J(l,m)^*\right)^{-1},
\end{equation}
where each term is a $2\times2$ complex matrix, with $J$ the voltage beam and $^*$ denoting the conjugate transpose. Because $J$ is complex and non-diagonal, all four combinations of the cross-correlated polarisations\footnote{The instrumental polarisations of the MWA are informally often referred to as $X$ and $Y$, mainly because many software treats the two polarisations as such. However, this is somewhat confusing because they are not necessarily orthogonal.} $p$ and $q$ including the imaginary part of the $pq$ and $qp$ are required to calculate any of the Stokes parameters. The $pp$ and $qq$ have no imaginary part due to the $uv$-symmetry. To make proper MWA beam correction possible, \textsc{wsclean} can output both real $pp$, $pq$, $qp$, $qq$ images and imaginary $pq$ and $qp$ images. This requires four runs of the algorithm. Once these images are created, a separate program is used to perform the correction of Eq.~\eqref{eq:beam-correction}. The method is demonstrated in Fig.~\ref{fig:pulsar-stokes-iv}, which displays a detection of pulsar J0437-4715 in Stokes V. The image was made with an initial pipeline for the MWA radio-sky monitor project that uses the MWA to search for transient and variable sources. The pipeline includes \textsc{wsclean} to make the Stokes-parameter images.

@@

\subsection{Decreasing inversion resolution} \label{sec:decreasing-inversion-resolution}
Typically, for cleaning it is desired that images have a pixel size (side length) at least five times smaller than the synthesised-beam width. This improves the accuracy of the clean algorithm, because the positions of image maxima will be closer to the actual source positions. This factor of five is normally taken into account in the overall imaging resolution, i.e., the image size is increased during inversion. For the $w$-projection method, the cost of inversion with an increased resolution is small, because it does not increase the size of the $w$-kernels. It will affect the inverse FFT, but the relative cost of this step is negligible in the total cost. The $w$-stacking method is however significantly affected by the image resolution, because it performs many inverse FFTs.

The spatial frequencies in the output image are band-limited by the synthesised beam. Therefore, it is not required to perform the inversion at higher resolution. As long as the $uv$-plane is sampled with at least the Nyquist frequency, a high resolution image can be perfectly reconstructed from an inversion at lower resolution. After an enlarged image has been cleaned, a model is created at the same (high) resolution. Because the model consists of delta functions, the spatial frequencies in the model are not band-limited. Consequently, decimating the model image will remove information from the model. However, only the low spatial frequencies of this image will be used in the prediction. Therefore, inversion at higher resolution is not required: as long as the decimation does not modify the low-frequency components, the output of the prediction step will not change. This requires low-pass filtering of the model before decimation.

We have implemented an option in \textsc{wsclean} to automatically decrease the inversion resolution to the Nyquist limit,
\begin{equation}\label{eq:nyquist-resolution}
 N_\textrm{Nyquist} = 2 \frac{N_\textrm{pix} S_\textrm{pix}}{S_\textrm{synth beam}},
\end{equation}
where $N_\textrm{Nyquist}$ is the resolution in pixels used during inversion, $N_\textrm{pix}$ is the requested number of pixels of the image along one side, $S_\textrm{pix}$ is the requested angular pixel scale and $S_\textrm{synth beam}$ is the minimum angular size of the synthesised beam. Before cleaning, the low-resolution image is interpolated with a procedure consisting of: i) a low-resolution FFT; ii) zero padding; and iii) a high-resolution inverse FFT. Such interpolation assumes the image to be periodic, but this is already assumed during the inversion process. After cleaning, the model is decimated by the inverse procedure, consisting of: i) a high-resolution FFT; ii) truncation; and iii) a low-resolution inverse FFT. With common settings, this procedure decreases the inversion imaging resolution by a factor of 2.5. Without further correction, such a decrease would lower the positional accuracy with which visibilities are gridded onto the $uv$-plane. This can be corrected by increasing the oversampling rate, which has almost no effect on performance.

Recreating Fig.~\ref{fig:vela-projection-example} using both the snapshot method and this optimisation lowers the computational cost from 41~min to 24~min. Of the 24~min, 77\% is spent on cleaning approximately 100,000 components. The outputs with and without rescaling do not visibly differ, but have an RMS difference of 20 mJy/beam, while the total RMS is 227 mJy/beam. Because the difference is mostly noise like, the difference could be caused by the non-linear behaviour of clean.
